\documentclass[fleqn]{article}

\usepackage{mydefs}
\usepackage{notes}
\usepackage{url}

\begin{document}
\lecture{CS4400}{Problem Set 4}{Alex Clemmer, u0458675}

% IF YOU ARE USING THIS .TEX FILE AS A TEMPLATE, PLEASE REPLACE
% "CS5350, Fall 2011" WITH YOUR NAME AND UID.

\section{3.62}

Please note the following information on your assignment:

\bee
\i What is the value of M?

\begin{solution}
13
\end{solution}

\i What registers hold program values \textit{i} and \textit{j}?

\begin{solution}
\texttt{\%ebx} and \texttt{\%ecx}
\end{solution}

\i Write a C code version of \texttt{transpose} that makes use of the optimizations that occur in this loop. Use parameter \texttt{M} in your code rather than numeric constants.

\begin{solution}
\end{solution}

\ene

\section{3.63}

Answer the following questions in 50-100 words each:
\bee
\i What is the difference between supervised, unsupervised and reinforcement learning?
% \begin{solution}
%   Blah..
% \end{solution}
\i List at least two real-world problems (other than those discussed in class) for each of the categories:
  supervised, unsupervised and reinforcement. For one of the supervised problems, what is the form
  of the output and what might be a reasonably input? For the reinforcement learning problem, what
  might be the state space, action space and reward function?
% \begin{solution}
%   Blah..
% \end{solution}
\i What is overfitting? What is underfitting? How is overfitting controlled?
% \begin{solution}
%   Blah..
% \end{solution}
\i Getting labeled data for supervised learning is expensive. What are some of the 
   approaches to deal with this problem?
% \begin{solution}
%   Blah..
% \end{solution}

\ene


\end{document}
