\documentclass{letter}
\signature{UgSAC}
\address{50 S. Central Campus Drive, Room 3190 \\ Salt Lake City, UT 84112}
\begin{document}
\begin{letter}{}
\opening{To Whom It May Concern:}

UgSAC would like to express unanimous and uncontested support for Matt Might as he makes his way through RPT this year. It is UgSAC's position

There are two things that really stand out to UgSAC in Might's portfolio. The first is Might's exceptional record of identifying and fostering the development of talented students. If you have read through his portfolio, you probably are already aware of the facts, but they bear repeating: one of Might's undergraduate students (David Darais) has gone on to the PhD program at Harvard, and another (Phillip Mates) was an intern at Google Summer of 2011. More critically, when we spoke to them, they both credited Might's leadership, expertise, and encouragement as critical points of inflection in their development.

This type of dedication is something that is easy enough to overlook or ignore for professors, and Might's devotion here is one thing that truly sets him apart from other professors, at least from the perspective of the undergraduate population. In UgSAC's opinion, it should be remembered that great members of the CS community are rarely simply born; rather, they are usually the product of some combination of talent and excellent mentorship. Great mentors are extremely rare, and we urge the committee to remember this as they make their decision.

The second thing that stands out is Might's exceptional contribution to the canon of undergraduate education here at Utah. We are speaking specifically about the undergraduate compilers course, which UgSAC views as a profound enrichment of the undergrad curriculum. One interesting thing about the class is that students are basically expected to work alone on a compiler for a mostly-inclusive subset of Python 3. This is interesting not only because it is notoriously difficult, but also because when you complete it you can list it on your resume---which is an explicit goal of the class.



\closing{Again, thank you,}

\end{letter}
\end{document}