\documentclass[ps, letterpaper]{cs121}
\usepackage{amsmath, amsfonts}

\setlength{\parindent}{0in}
\setlength\parskip{2mm}

\begin{document}
\header{9}{Tuesday, December 7, 2010}{NO LATE DAYS}{20}

% !!!!! Write your name here !!!!! %
\vspace{3mm}
\begin{center} \bf{Name} \end{center}
\studentName{!!! Your Name Here !!!}
\collaboratorNames{!!! Collaborators' names here !!!!}

\newcommand{\<}{\langle}
\renewcommand{\>}{\rangle}

\renewcommand{\P}{\mathrm{P}}
\renewcommand{\NP}{\mathrm{NP}}
\newcommand{\NST}{\mathrm{NST}}
\newcommand{\CoNP}{\mathrm{\text{Co-}NP}}
\newcommand{\NPC}{\mathrm{NP}\text{-complete}}
\newcommand{\PSPACE}{\mathrm{PSPACE}}
\newcommand{\NPSPACE}{\mathrm{NPSPACE}}
\newcommand{\DSPACE}{\mathrm{DSPACE}}
\newcommand{\DoubleSAT}{\textsc{Double-Sat}}
\newcommand{\Primes}{\textsc{Primes}}
%\newcommand{\SATsearch}{\textsc{SATSearch}}
\newcommand{\abs}[1]{\lvert#1\rvert}

\setlength\parskip{2mm}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% PROBLEM 1 - 
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\problem{5 + 5} %2005
\subproblem
Let $\DoubleSAT= \{\<\phi\>: \text{$\phi$ is a boolean formula with at least two satisfying assignments }\}$. Show that $\DoubleSAT$ is $\NPC$.

\subproblem
Let $k$ be a natural number, and let ${\cal C}$ be a finite collection of finite sets. Then ${\cal C}$ is said to be a $k$-{\sc Hitting Set} if there is a set $H$ of size at most $k$ that intersects every member of ${\cal C}$. That is, a finite collection $C$ of finite sets is  a $k$-{\sc Hitting Set} if there exists a set $H$ of size at most $k$ such that $S\cap H\neq\emptyset$ for every $S \in {\cal C}$. Prove deciding $k$-{\sc Hitting Set} is $\NP$-complete.  [\emph{Hint}: Reduce from {\sc Vertex Cover}.]

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% PROBLEM 2 - Prove NP-Complete harder
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\problem{5+5+2} %2007
Consider the following two languages:

$$A=\{\left<D\right>\text{ : $D$ is a DFA and $D$ accepts at least one string}\}$$
$$B=\{\left<D_1,D_2,...,D_k\right>\text{ : Each $D_i$ is a DFA and all of the $D_i$ accept at least one string in common} \}$$

\subproblem Show that $A \in \P$.

\subproblem Show that $B$ is $\NP$-complete by giving a reduction from 3-SAT to $B$.

\subproblem We can convert an instance of $B$ into an instance of $A$ by applying the product construction (See p.45-46 of {\em Sipser}) $k-1$ times in succession. Does this show that $\P=\NP$? Why or why not?

\pagebreak
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% PROBLEM 3 - Co-NP
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\problem{5+2} %2007

Recall that $\CoNP = \{ L: \b{L} \in \NP \}$.  It is unknown whether or not $\NP = \CoNP$. Note that $\NP = \CoNP$ if and only if $\NP$ is closed under complement.

\subproblem
Prove that if $\NP \neq \CoNP$, then $\P \neq \NP$. (Aside: Nonetheless, we can't rule out the possibility that $\NP = \CoNP$ and yet $\P \neq \NP$.)

\subproblem
To prove that $\NP = \CoNP$, it would suffice to show that for every $L \in \NP$, $\b{L} \in \NP$.  Suppose $L \in \NP$.  Then there exists a nondeterministic Turing machine $M$ that decides $L$ in polynomial time.  Consider the new Turing machine $M^\prime$, which is identical to $M$ except that its accept and  reject states are reversed.

What language does $M^\prime$ decide?  Explain \emph{briefly}.


%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% PROBLEM 4 - NP-Search
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\problem{5+5+10} 

A {\em search problem} is a mapping $S : \Sigma^* \rightarrow P(\Delta^*)$ from strings (``instances'') to sets of strings (``valid solutions''). An algorithm $M$ {\em solves} a search problem $S$ if for every input $x\in \Sigma^*$ such that $S(x)\neq \emptyset$, $M$ outputs some solution in $S(x)$.

An {\em $\NP$ search problem} is one for which there exists a polynomial $p$ and a polynomial-time algorithm $V$ such that for every $x$ and $y$

\begin{enumerate}\addtolength{\itemsep}{-0.3\baselineskip} 
\item $y\in S(x) \Rightarrow |y|\leq p(|x|)$, and 
\item $y\in S(x) \Leftrightarrow V$ accepts $\langle x,y\rangle$.
\end{enumerate}

Informally, these conditions say that valid solutions are short and can be verified efficiently.

\subproblem Show that the $\SATsearch$ problem ``given a satisfiable boolean formula $\varphi$, find a satisfying assignment'' is an $\NP$ search problem

\subproblem Prove that $\SAT\in\P$ if and only if the $\SATsearch$ problem can be solved in polynomial time.

\subproblem Prove that $\P=\NP$ if and only if every $\NP$ search problem can be solved in polynomial time.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% PROBLEM 5 - Space and (nondeterministic) time
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\problem{BONUS +1}
A language is in  $\NST(S,T)$ if it is accepted by a nondeterministic Turing Machine that runs simultaneously in space $S(n)$ and time $T(n)$. Show that $\NST(S(n),T(n)) \subseteq \DSPACE(S(n) \log T(n))$. Then show that actually $\NST(S(n),T(n)) \subseteq \DSPACE(S(n) \log (T(n)/S(n)))$ (an improvement, for example, if $T(n) = S(n)\log S(n)$).

\end{document}